Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-11-25 manuscript.qmd
remove comments and add trial citation

<!-- However, for early T-category scenarios with midline extension, the model does seem to overestimate contralateral LNL II and III invovlement. This likely stems from the the small sample size of this relatively rare scenario as hinted at by the wide beta posteriors. -->

@@

<!-- However, there are some discrepancies in the model's prediction: It cannot quite capture the correlation of ipsi- and contralateral involvement that is seen in the data. This is notable for example in the bottom right subplot where the prevalence is overestimated for the scenario of an (almost) healthy ipsilateral neck and underestimated when at least LNL III shows metastases. One reason for this may be the width of the time-prior: If it was wider, the posterior over the diagnosis time given the ipsilateral involvement could more flexibly shift to reflect a more severe contralateral involvement. -->

<!-- NOTE: This last sentence is purely speculative! We need to discuss this. -->

@@

<!-- @Fig-model-risks shows this predicted risk of occult disease in the contralateral LNLs II, III, and IV for interesting combinations of these three risk factors. This risk is computed *given* a CT diagnosis assessing the per-LNL clinical involvement for which we assumed a sensitivity of 81% and a specificity of 76%. For the involvement in the LNls III and IV contralaterally, we also cosidered how the risk would be affected if the upstream LNL was confirmed to be metastatic using fine needle aspiration (FNA) with a specificity of 98% and a sensitivity of 80% [@debondt_detection_2007]. In the figure legends, this is noted e.g. for the orange historgam over the risk of contralateral LNL IV involvement as `c:III FNA+`, meaning that the contralateral LNL III was pathologically confirmed to harbor metastases via FNA.-->

@@

The predictions of this model are being used to inform a clinical trial on volume deescalation at the University Hospital Zurich [clinicaltrials.gov, NCT06563362 (DeEscO)<!-- TODO: add -->]. When accepting a 5% risk of occult disease in any given LNL, the model suggests the following contralateral elective irradiation (assuming the respective LNL is clinically healthy):

>>>>

@@

@@

@@

The predictions of this model are being used to inform a clinical trial on volume deescalation at the University Hospital Zurich [@universityofzurich_personalized_2024]. When accepting a 5% risk of occult disease in any given LNL, the model suggests the following contralateral elective irradiation (assuming the respective LNL is clinically healthy):
